---
title: "Working with XLSForm in novaXLSflow"
author: "Your Name"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xlsform-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
# This code block won't run until package is built,
# we assume you have the functions from previous steps
# installed and available in the novaXLSflow package.

```

# Introduction

This vignette demonstrates how to parse an XLSForm, convert it into a flowchart structure,
save it as JSON, and visualize it using novaXLSflow.

## Installation

If you haven’t already installed novaXLSflow, do so (adjust the GitHub username and repo as needed):

```{r}
# install.packages("devtools") # if needed
# devtools::install_github("yourusername/novaXLSflow")
library(novaXLSflow)
```


Parsing an XLSForm

We placed an example XLSForm in inst/extdata/example_xlsform.xlsx within the package.
You can access it using system.file():

```{r}
xlsform_path <- system.file("extdata", "example_xlsform.xlsx", package = "novaXLSflow")
flowchart_data <- parse_xlsform(xlsform_path)

# Examine the structure
str(flowchart_data)
```


The flowchart_data is a list with nodes and edges that represent the form questions and their logical relationships.
Saving as JSON

You can save the parsed structure to JSON for further inspection or integration with other tools:

```{r}
save_flowchart_json(flowchart_data, "flowchart.json")
```


This will create a flowchart.json file in your working directory with a machine-readable structure of the form.
Visualizing the Flow

Use visualize_flow() to create a Graphviz-based flowchart of the XLSForm logic:

```{r}
graph <- visualize_flow(flowchart_data)
graph
```


When viewed in RStudio or a browser, this will display a diagram showing each question node and edges indicating relevance conditions.

